---
title: "R Notebook"
output: html_notebook
---

**Exercise 1. Additive Decomposition: Beer Production Data**


```{r}
#rm(list = ls())

library(fpp3)
library(dbplyr)

aus_production

class(aus_production$Quarter) # "vctrs_vctr"
class(aus_production$Beer) # numeric
              

aus_production |>
  filter_index("1992 Q1" ~ "2000 Q4") |>
  select(Quarter, Beer)

```

```{r}

# Extract quarterly beer production from 1992 Q1 to 2000 : (way 1)
beer_prod <- aus_production %>%
  filter(Quarter >= yearquarter("1992 Q1"),
         Quarter <= yearquarter("2000 Q4")) |>
  select(Quarter, Beer)

```

1. Get the data

```{r}
#equivalently 

# Extract quarterly beer production from 1992 Q1 to 2000 : (way 2)

beer_prod <- aus_production |>
  filter_index("1992 Q1" ~ "2000 Q4") |>
  autoplot(Beer) +
  labs(
    title = "Australian beer production",
    x = "Quarter",
    y = "Beer production"
  )

beer_prod

# Careful after autoplot(Beer), beer_prod is now an object of class ggplot not a tsibble anymore
```

2. Estimate the trend using LOESS

```{r}
beer_prod <- aus_production |>
  filter_index("1992 Q1" ~ "2000 Q4") |>
  select(Quarter, Beer)


# We need to recreate beer_prod from aus_production to get an object of class tsibble

beer <- beer_prod |>
  mutate(
    time = row_number()
  )

loess_fit <- loess(Beer ~ time, data = beer) # ~ option + N + espace

beer <- beer |>
  mutate(
    Trend = predict(loess_fit)
  )

beer
```

3. Detrend the data

```{r}
beer <- beer |>
  mutate(
    Detrended = Beer - Trend,
    Quarter_num = quarter(Quarter)
  )

beer

```

4. Calculate seasonal effects

```{r}
seasonal <- beer |>
   as_tibble() |>
   group_by(Quarter_num) |>
   summarise(
    Seasonal = mean(Detrended, na.rm = TRUE),
    .groups ='drop'
  )

seasonal
beer


```

5. Add the seasonal component back to the data

```{r}

beer <- beer |>
  mutate(
    Quarter_num = quarter(Quarter)
  ) |>
  left_join(seasonal, by = "Quarter_num")

beer
```

6. Calculate the remainder

```{r}

beer <- beer |>
  mutate(
    Remainder = Beer - Trend - Seasonal
  )
beer
```

7. Plot the components

```{r}

beer |>
  autoplot(Beer) +
  geom_line(aes(y = Trend), linewidth = 1)

```

Plot the individual components separately:

```{r}

beer |>
  ggplot(aes(x = Quarter, y = Beer)) +
  geom_line() +
  geom_line(aes(y = Trend), linewidth = 1, col = 'red') 
  

```

For the seasonal component:

```{r}
beer |>
  ggplot(aes(x = Quarter, y = Seasonal)) +
  geom_line() +
  labs(
    title = 'Estimated seasonal component',
    x = 'Quarter',
    y= 'Seasonality'
  )
  
```

And the remainder:

```{r}

beer |>
  ggplot(aes(x = Quarter, y = Remainder)) +
  geom_line() +
  labs(
    title = 'Estimated remainder component',
    x = 'Quarter',
    y= 'Remainder'
  )

```

Perform an STL decomposition using STL()

```{r}

beer_stl <- beer |>
  model(
    STL(Beer ~ trend(window = 7) + 
          season(window = 'periodic'))
  )
```

Plotting STL decomposition

```{r}

components(beer_stl) |>
  autoplot()

```
