---
title: "R Notebook"
output: html_notebook
---

**Exercise 2. Synthetic Additive Time Series**

Generate a quarterly additive series with linear trend, seasonal component, and Gaussian noise:

```{r}
library(fpp3)

n <- 50
sigma <- 2
epsilon <- rnorm(n, mean = 0, sd = sigma)

# Linear trend
T <- 0.05 * (1:n)

# Quarterly seasonal component
S <- rep(c(-2, 1, 3, -2), length.out = n)
  

# Additive time series
Y <- T + S + epsilon

```

Plot the generated series:

```{r}

plot(Y, type = "l",
     main = "Quarterly Additive Time Series",
     xlab = "Time",
     ylab = "Y")

```
Perform a manual decomposition using LOESS for the trend and average detrended values per
quarter for seasonality.

Estimating the trend:

```{r}

time <- 1:n

quarter <- rep(1:4, length.out = n)

loess_fit <- loess(Y ~ time) 

T_hat <- predict(loess_fit)
T_hat

plot(time, Y, type = "l",
     main = "Observed Series and LOESS Trend",
     xlab = "Time",
     ylab = "Y")

lines(time, T_hat, lwd = 2)

ggplot() +
  aes(x = time, y = T_hat) +
  geom_line() +
  labs(
    title = "Estimated trend",
    x = "Time",
    y = 'Trend'
  )


```

Detrend the data:

```{r}
Detrend = Y - T_hat
Detrend
```

Average the detrended values by quarter:

```{r}

seasonal_means <- tapply(Detrend, quarter, mean)
seasonal_means
sum(seasonal_means) # ≠ 0, so we must center it

```

Center the seasonal component:

```{r}

s_hat_quarter <- seasonal_means - mean(seasonal_means)
s_hat_quarter
sum(s_hat_quarter) # = 0

S_hat <- s_hat_quarter[quarter]
S_hat

ggplot() +
  aes(x = time, y = S_hat) +
  geom_line() +
  labs(
    title = "Estimated seasonal component",
    x = "Time",
    y = 'Seasonality'
  )



```

Estimate the noise/remainder term:

```{r}

epsilon_hat = Y - T_hat - S_hat
epsilon_hat

ggplot() +
  aes(x = time, y = epsilon_hat) +
  geom_line() +
  labs(
    title = "Estimated remainder component",
    x = "Time",
    y = 'Remainder'
  )


```

Perform STL decomposition:

```{r}

Y_ts <- ts(Y, frequency = 4)
Y_stl <- stl(Y_ts, s.window = "periodic")
plot(Y_stl)

```

